---
title: "Getting started with maudr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with maudr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## Overview

**maudr** generates individualised enzyme kinetics datasets for undergraduate teaching. Each student receives a unique Excel file containing simulated absorbance data for an inhibited and non-inhibited enzyme reaction. Instructors can then produce per-student answer PDFs with plots and estimated kinetic parameters.

The package ships with demo data — a list of four imaginary students and kinetic parameters for yeast alcohol dehydrogenase (ADH) — so you can run the full workflow without preparing any input files.

The four main functions must be used in order:

1. `initialiseProject()` — set up the folder structure
2. `assignReactions()` — assign each student a substrate and inhibitor
3. `generateAssignments()` — produce per-student Excel data files
4. `generateAnswers()` — produce per-student answer PDFs

---

## Installation

**maudr** is available from GitHub only:

```{r install}
install.packages("remotes")
remotes::install_github("jarekbryk/maudr")
```

```{r library}
library(maudr)
```

---

## Step 1: Initialise the project

`initialiseProject()` creates a self-contained folder structure:

```
maudr_assignments/
  data/
  output/
    assignments_output/
    answers_output/
```

```{r init}
initialiseProject(path = "~/Desktop")
```

By default the top-level folder is named `maudr_assignments` and is created in the current working directory. Both can be changed:

```{r init-custom}
initialiseProject(path = "~/Documents", name = "enzyme_project")
```

After `initialiseProject()` is called, **the project path is registered automatically** and all subsequent functions will find it without you needing to pass `project_path` again. You can still supply `project_path` explicitly if you want to override this.

---

## Step 2: Assign reactions to students

`assignReactions()` randomly assigns each student a substrate and an inhibitor, then writes a timestamped metadata file to `output/assignments_output/`.

### Demo mode

If no input files are provided, the function uses the bundled demo student list (four imaginary students) and the default ADH parameters:

```{r assign-demo}
setup <- assignReactions()
```

### With your own files

Supply a student list and (optionally) a custom enzyme parameters file:

```{r assign-custom}
setup <- assignReactions(
  student_file = "~/Documents/students.xlsx",
  enzyme_file  = "~/Documents/my_enzyme_params.xlsx"
)
```

The student list must contain the columns `student_no`, `first_name`, and `surname`. The enzyme parameters file must contain `rxn_substrate`, `Kcat`, `Km`, `Vmax`, `enzyme_conc`, and `inhibition_actual`. Both input files are automatically copied into the project's `data/` folder.

The function returns a list with two elements:

- `$timestamp` — a string (e.g. `"2025-08-22_0848"`) used to identify this run in later steps
- `$metadata` — the full student–reaction assignment table

---

## Step 3: Generate student assignment files

`generateAssignments()` reads the metadata from Step 2 and writes one Excel file per student to `output/assignments_output/<timestamp>/`. Each file contains simulated absorbance vs time data across a range of substrate concentrations, for both the inhibited and uninhibited condition.

```{r gen-assign}
generateAssignments(run_timestamp = setup$timestamp)
```

A small, reproducible amount of random noise is added to each student's data by default (`use_jitter = TRUE`), simulating experimental variation. You can adjust the noise level or disable it:

```{r gen-assign-nojitter}
generateAssignments(run_timestamp = setup$timestamp, use_jitter = FALSE)
generateAssignments(run_timestamp = setup$timestamp, jitter_sd = 0.05)
```

---

## Step 4: Generate answer PDFs

`generateAnswers()` reads each student's Excel file and produces:

- Absorbance vs time plot (faceted by condition, coloured by substrate concentration)
- Michaelis-Menten plot (both conditions on one plot, with optional fitted curve)
- Lineweaver-Burk plot (both conditions, with fitted regression lines and equations)
- A summary table of estimated K~m~ and V~max~ per condition

```{r gen-answers}
generateAnswers(run_timestamp = setup$timestamp)
```

By default a single combined PDF (`answers_all_students.pdf`) is written to `output/answers_output/<timestamp>/`. Use `output_files` to control this:

```{r gen-answers-opts}
# One PDF per student
generateAnswers(run_timestamp = setup$timestamp, output_files = "separate")

# Both combined and individual files
generateAnswers(run_timestamp = setup$timestamp, output_files = "both")
```

---

## Full demo workflow

The complete workflow in demo mode, using the bundled data:

```{r full-demo}
library(maudr)

initialiseProject()

setup <- assignReactions()

generateAssignments(run_timestamp = setup$timestamp)

generateAnswers(run_timestamp = setup$timestamp, output_files = "both")
```

---

## Using your own enzyme

To use an enzyme other than ADH, provide a custom `enzyme_file` with the required columns. The `inhibition_actual` column must contain at least one non-`"no_inhibition"` value; `"no_inhibition"` rows are used internally to represent the uninhibited condition and are generated automatically.

---

## Reproducibility

`assignReactions()` accepts a `seed` argument (default `1234`) which controls both the random assignment of students to conditions and the noise added in `generateAssignments()`. Setting the same seed on the same student list will always produce identical assignments and data files.
